### 2024-06-20 manuscript.tex
Auto title date
Writing introduction

@@

\begin{center} \Large \sffamily\textbf{Using ship observations to assess Southern Ocean clouds in a storm-resolving general circulation model ICON}\\[0.4cm]
\large Peter Kuma$^\mathsf{1,}$*, Frida A.-M. Bender$^\mathsf{1}$, Adrian J. McDonald$^\mathsf{2}$ and Thorsten Mauritsen$^\mathsf{1}$\\[0.4cm]
\small
$^\mathsf{1}$Department of Meteorology (MISU) and Bolin Centre for Climate Research, Stockholm University, Stockholm SE-106 91, Sweden\\[0.2cm]
$^\mathsf{2}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, New Zealand\\[0.2cm]
*Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\large 11 June 2024\\[0.4cm]

@@

Currently, a new generation of km-scale resolution global climate models are in development as the forthcoming phase of climate modelling. One such model is a 5-km version of the Icosahedral Nonhydrostatic Weather and Climate Model (ICON) developed jointly by Deutscher Wetterdienst (DWD) and the Max-Planck-Institute for Meteorology (MPI-M). Because of the high resolution, most parametrisations, such as that of convection and clouds, can be avoided.

Previous studies have identified substantial large-scale biases in model clouds over the Southern Ocean (SO), affecting sea surface temperature and the Earth's albedo overall. Our aim is to quantify how well the high-resolution ICON model is simulating clouds in this region, particularly in light of the fact that subgrid-scale clouds are not parametrised in this model. This region is mostly dominated by boundary layer clouds generated by shallow convection, and these are problematic to observe by spaceborne lidar and radars, which are affected by attenuation by overlapping and thick clouds and ground clutter, respectively. Therefore, we choose to use a large set of ship-based observations conducted with ceilometers and lidars on board of RV \textit{Polarstern} and other voyages. Altogether, we analyse over 2000 days (about 6 years) of data from 31 voyages and 1 sub-antarctic station covering diverse longitudes of the SO. To achieve a like-for-like comparison with the model, we use a ground-based lidar simulator called the Automatic Lidar and Ceilometer Framework (ALCF). We contrast the results with the ECMWF Reanalysis 5 (ERA5) and the Modern-Era Retrospective analysis for Research and Applications, Version 2 (MERRA-2).

@@

Increasing climate model resolution is a common way of improving model
accuracy \citep{mauritsen2022}. This is of course limited by the available
computational power and a trade-off with increasing parametrisation complexity.
Current computational availability is progressing to enable km-scale Earth
system models (ESMs) and coupled atmophsere--ocean general circulation models
(AOGCMs) in research conditions today and operationally in the forthcoming
years. Therefore, it represents a natural advacement in climate modeling.
Representative of this efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019}, which is an intercomparison of global storm-resolving
models.
% Grundner (2023), Data-Driven Cloud Cover Parameterizations for the ICON Earth System Model Using Deep Learning and Symbolic Regression

The NextGEMS project is focused on the research and
development of high resolution (km-scale) global storm-resolving climate models
(GSRMs), which are planned to be the next generation of operational climate
models. Because of the high resolution, some previously parameterised physical
processes can be resolved on the model grid. Thus, such parameterisations and
related uncertainties can be eliminated.

The main aim of our research is to evaluate a high-resolution version of the
Icosahedral Nonhydrostatic Weather and Climate Model (ICON), developed by
NextGEMS jointly with Deutscher Wetterdienst (DWD) and the Max-Planck-Institute
for Meteorology (MPI-M). Previous studies have identified substantial
large-scale biases in climate model clouds over the Southern Ocean (SO), affecting
sea surface temperature and the Earth’s albedo overall. Our aim is to quantify
how well the high-resolution ICON model is simulating clouds in this region,
particularly in light of the fact that subgrid-scale clouds are not
parameterised in this model. This region is mostly dominated by boundary layer
clouds generated by shallow convection, and these are problematic to observe by
spaceborne lidar and radars, which are affected by attenuation by overlapping
and thick clouds and ground clutter, respectively. Therefore, we chose to use
a large set of ship-based observations conducted with ceilometers and lidars on
board of the RV \emph{Polarstern} and other voyages and stations.

Marine clouds have disproportionate effect on TOA shortwave radiation due to
the relatively low albedo of the sea surface. The longitudinal symmetry of the
SO means that model cloud biases tend to be similar across
longitudes.

@@

\begin{figure}
\centering
\includegraphics[width=0.65\textwidth]{img/map.png}
\caption{A map showing the tracks of 31 voyages of RV \emph{Polarstern}, RSV \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer} and HMNZS \emph{Wellington} and one sub-Antractic station (Macquarie Island) analysed here. The tracks cover Antarctic sectors south of South Americ, the Atlantic Ocean, Africa, Australia and New Zealand in the years 2010 to 2021.}
\label{fig:map}
\end{figure}

@@

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, RSV
\emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer},
HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island) in the
SO south of 40°S between year 2010 and 2021 (Fig. \ref{fig:map} and
Table \ref{tab:voyages}). Altogether, the voyages and the station comprised
2421 days of data south of 40$^\circ$S, but the availability of ceilometer data
was smaller.

@@

A coupled storm-resolving version of the ICON model is in development by the
NextGEMS project \citep{hohenegger2023}. NextGEMS has so far produced four
cycles of model runs. In our research, we used a Cycle 3 run ngc3028 produced
in 2023 \citep{nextgems2023a,nextgems2023b}. We could not use Cycle 4 due to unavailability of
the necessary variables. The horizontal resolution of ngc3028 is about 5 km.
Unlike current general circulation models (GCMs), the strom-resolving version
of ICON does not use convective and cloud parameterisation but relies on
explicit simulation of convection and clouds on the model grid. While this
makes the code development simpler without having to rely on uncertain
parameterisations, it can miss smaller-scale clouds below the grid resolution.
Turbulence and cloud microphysics are still parameterised in this model. In our
analysis, we used four years of coupled simulation output between years 2021
and 2024 (inclusive). Because the model is free-running, weather and climate
oscillations (such as the El Niño--Southern Oscillation) are not expected to be
equivalent to reality at the same time and place. To compare with observations
taken during different years (2010--2021), we compared the model output with
observations at the same time of year and geographical location.

@@

>>>>

% Grundner (2023), Data-Driven Cloud Cover Parameterizations for the ICON Earth System Model Using Deep Learning and Symbolic Regression [thesis]
% Satoh et al. (2019), Global Cloud-Resolving Models

@@

\begin{center} \Large \sffamily\textbf{Using ship observations to assess Southern Ocean clouds in a storm-resolving general circulation model ICON}\\[0.4cm]
\large Peter Kuma$^\mathsf{1,}$*, Frida A.-M. Bender$^\mathsf{1}$, Adrian J. McDonald$^\mathsf{2}$ and Thorsten Mauritsen$^\mathsf{1}$\\[0.4cm]
\small
$^\mathsf{1}$Department of Meteorology (MISU) and Bolin Centre for Climate Research, Stockholm University, Stockholm SE-106 91, Sweden\\[0.2cm]
$^\mathsf{2}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, New Zealand\\[0.2cm]
*Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\large \today\\[0.4cm]

@@

Currently, a new generation of km-scale (also called `k-scale') resolution
global climate models are in development as the forthcoming phase of climate
modelling. One such model is a 5-km version of the Icosahedral Nonhydrostatic
Weather and Climate Model (ICON) developed jointly by Deutscher Wetterdienst
(DWD) and the Max-Planck-Institute for Meteorology (MPI-M). Because of the high
resolution, most parametrisations, such as that of convection and clouds, can
be avoided.

Previous studies have identified substantial large-scale biases in model clouds
over the Southern Ocean (SO), affecting sea surface temperature and the Earth's
albedo overall. Our aim is to quantify how well the high-resolution ICON model
is simulating clouds in this region, particularly in light of the fact that
subgrid-scale clouds are not parametrised in this model. This region is mostly
dominated by boundary layer clouds generated by shallow convection, and these
are problematic to observe by spaceborne lidar and radars, which are affected
by attenuation by overlapping and thick clouds and ground clutter,
respectively. Therefore, we choose to use a large set of ship-based
observations conducted with ceilometers and lidars on board of Research Vessel
(RV) \textit{Polarstern} and other voyages. Altogether, we analyse over 2000
days (about 6 years) of data from 31 voyages and 1 sub-antarctic station
covering diverse longitudes of the SO. To achieve a like-for-like comparison
with the model, we use a ground-based lidar simulator called the Automatic
Lidar and Ceilometer Framework (ALCF). We contrast the results with the ECMWF
Reanalysis 5 (ERA5) and the Modern-Era Retrospective analysis for Research and
Applications, Version 2 (MERRA-2).

@@

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a tradeoff with increasing parametrisation complexity.  Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPGPUs) is progressing to enable km-scale Earth
system models (ESMs) and coupled atmosphsere--ocean general circulation models
(AOGCMs) in research conditions today and operationally in the forthcoming
years. Therefore, it represents a natural advancement in climate modeling.
Global strom-resolving models (GSRMs) are emerging as a new front in the
development of high-resolution global climate models, with horizontal grid
resolutions of about 2--8 km \citep{satoh2019,stevens2019}. This is enough to
resolve mesoscale convective storms, but smaller-scale convective plumes and
smaller-scale cloud structure remain unresolved. At a 5-km scale,
non-hydrostatic processes also become important, and for this reason such
models are generally non-hydrostatic. The terms global cloud-resolving models
(GCRMs), or global convection-permitting/-resolving models (GCPMs) are also
sometimes used interchangeably with GSRMs but imply that clouds or convection
are resolved explicitly, which is not entirely true for GSRMs, as this would
require an even higher horizontal resolution \citep{satoh2019}.  Representative
of these efforts is the DYnamics of the Atmospheric general circulation Modeled
On Non-hydrostatic Domains (DYAMOND) project \citep{stevens2019,dyamond}, which
is an intercomparison of nine global storm-resolving models (GSRMs) on two
40-day time periods in summer (1 August -- 10 September 2016) and winter (20
January - 1 March 2020). A new one-year GSRM intercomparison is currently
proposed by \cite{takasuka2024}, with the hope of also evaluating the seasonal
cycle and large-scale circulation.  An alternative to using a
computantionally costly GSRM is to train an artificial neural network on GSRM
output and use it for subgrid-scale clouds, as done with the GSRM ICON by
\cite{grundner2022}; \cite{grundner2023}.

nextGEMS is an EU-funded project \citep{nextgems} focused on the research and
development of GSRMs at multiple modelling centres and universities in Europe.
The project also develops GSRM versions of the Icosahedral Nonhydrostatic
Weather and Climate Model (ICON), the Integrated Forecasting System (IFS), and
their ocean components at eddy-resolving resolutions: ICON-O coupled with ICON
and Finite-Element/volumE Sea ice-Ocean Model (FESOM) and Nucleus for European
Modelling of the Ocean (NEMO) coupled with IFS.  The project has so far
produced ICON and IFS simulations in three cycles (Cycle 1--3) and pre-final
simulation, with a final production simulation planned by the end of the
project. nextGEMS is not the only project developing GSRMs. Other GSRMs (or
GSRM versions of climate models) currently in development include:
Convection-Permitting Simulations With the E3SM Global Atmosphere Model
[SCREAM; \cite{caldwell2021}], Atmospheric Model [NICAM; \cite{satoh2008}],
Unified Model (UM), eXperimental System for High-resolution modeling for
Earth-to-Local Domain [X-SHiELD; \cite{shield}], Action de Recherche Petite
Echelle Grande Echelle-NonHydrostatic version [ARPEGE-NH;
\cite{bubnova1995,voldoire2017}], Finite-Volume Dynamical Core on the Cubed
Sphere [FV3, \cite{lin2004}], NASA Goddard Earth Observing System global
atmospheric model version 5 [GEOS5; \cite{putman2011}], Model for Prediction
Across Scales [MPAS; \cite{skamarock2012}], and System for Atmospheric Modeling
[SAM; \cite{khairoutdinov2003}].

Multiple cloud properties have an effect on the shortwave (SW) and longwave
(LW) radiation. On the first order, it is the total cloud fraction, cloud phase,
and the liquid and ice water path. These are influenced by the atmopheric
thermodynamics, convection and circulation, and indirect and direct effects
of aerosols. On smaller orders, it is the cloud droplet size distribution, ice
crystal habit, cloud lifetime, and direct radiative interaction with aerosols.
In the 6th phase of the Coupled Model Intercomparison Project [CMIP6;
\citep{eyring2016}], the cloud feedback has increased relative to CMIP5
\citep{zelinka2020}, also being one of the main reasons for the higher climate
sensitivity of CMIP6 models.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to reasons such as a lack of surface and in situ observations and
being a lower priority region for numerical weather prediction (NWP) and
climate model development because it is far away from populated areas of
interest.  Nevertheless, radiation biases and changes over an area of its size
have a very significant influence on the global climate, and the SO is an
important part of the global ocean conveyor belt.  Marine clouds have
disproportionate effect on top of atmosphere (TOA) SW radiation due to the
relatively low albedo of the sea surface.  The longitudinal symmetry of the SO
means that model cloud biases tend to be similar across longitudes.  Here, we
will conventionally refer to the SO as ocean regions south of 40$^\circ$S,
low-latitude SO as 40--55$^\circ$S and high-latitude SO as south of
55$^\circ$S.

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW CRE
bias is still positive in eight analysed CMIP6 models analysed by
\cite{schuddeboom2021} over high-latitude SO, whereas over low-latitude SO it
tends to be more neutral or negative in some models. Too much absorbed SW
radiation over the SO was also identified in the GSRM SCREAM
\cite{caldwell2021}. Compensating biases are possible, such as the `too few too
bright' cloud bias, characterised by too small cloud fraction and too large
cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a resonable SW radiation balance by
reflecting too much SW radiation by cloudy sky but has too little cloudy sky
area overall. A study by \cite{konsta2022} showed that this type bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \citep{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to the Moderate Resolution Imaging
Spectroradiometer (MODIS)] and overestimates downwelling TOA SW radiation on
the order of 10 Wm$^\mathrm{-2}$ in  [relative to the Clouds and the Earth’s
Radiant Energy System (CERES)] in the highest model resolution run (2.5 km).
\cite{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and total CF underestimation on the order of 10\%
over the entire 40--60$^\circ$S SO.

In general, SST biases in the SO can originate either in the atmosphere, caused
by too much shortwave heating of the surface, too little longwave cooling of
the surface, or in the ocean circulation.  Interactions of both are also
possible, due to, for example SST affecting clouds and clouds affecting the
surface radiation.  \cite{zhang2023} have shown that SST biases have improved
in CMIP6 compared to CMIP5 (relative to ERA5), with SST overall increasing in
the later CMIP phase. However, over the SO this resulted in an even higher
positive bias, especially in the Atlantic Ocean (AO) sector of the SO,
increasing by up to 1$^\circ$C.  \cite{luo2023} identified that the SO SST bias
in an ensemble of 18 CMIP6 models originates not from the surface heat and
radiation fluxes (using reanalyses as a reference), but from a warm bias in the
Northern Atlantic Deep Water.

The main aim of our research is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst (DWD) and the Max-Planck-Institute
for Meteorology (MPI-M). Previous studies have identified substantial
large-scale biases in climate model clouds over the Southern Ocean (SO),
affecting sea surface temperature and the Earth’s albedo overall. Our aim is to
quantify how well the high-resolution ICON model is simulating clouds in this
region, particularly in light of the fact that subgrid-scale clouds are not
parameterised in this model. This region is mostly dominated by boundary layer
clouds generated by shallow convection, and these are problematic to observe by
spaceborne lidar and radars, which are affected by attenuation by overlapping
and thick clouds and ground clutter, respectively.  Specifically, the radar on
CloudSat and lidar on CALIPSO are affected by the abovementioned issues,
resulting in a strong underestimation of cloud occurrence below 2 km relative to
ground-based lidar observations \citep{mcerlich2021}.  This, in turn, can lead to
systematic biases in low clouds in climate models, which are frequently
evaluated against CloudSat--CALIPSO products. Reanalyses can also suffer from
cloud biases, as these are usually parametrised in their atmospheric component,
and also in regions where input observations are sparse.  This makes them a
problematic reference for clouds over the SO, and any biases relative to a
reanalysis should be interpreted with caution. Instead, we chose to use a large
set of ship-based observations conducted with ceilometers and lidars on board
of the RV \emph{Polarstern} and other voyages and stations as a reference for
the model evaluation.

@@

\begin{figure}
\centering
\includegraphics[width=0.65\textwidth]{img/map.png}
\caption{
A map showing the tracks of 31 voyages of RV \emph{Polarstern}, RSV
\emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer}, and
HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island)
analysed here. The tracks cover Antarctic sectors south of South America, the
Atlantic Ocean, Africa, Australia, and New Zealand in the years 2010 to 2021.
The dashed line at 40$^\circ$S delineates the Southern Ocean area of our
analysis.
}
\label{fig:map}
\end{figure}

@@

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, RSV
\emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer}, Her
Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one sub-Antarctic
station (Macquarie Island) in the SO south of 40°S between year 2010 and 2021
(Fig. \ref{fig:map} and Table \ref{tab:voyages}). Altogether, the voyages and
the station comprised 2421 days of data south of 40$^\circ$S, but the
availability of ceilometer data was smaller.

@@

A coupled (atmosphere--ocean) GSRM version of the ICON model is in development
by the nextGEMS project \citep{hohenegger2023}. nextGEMS has so far produced
four cycles of model runs. In our research, we used a Cycle 3 run
\emph{ngc3028} produced in 2023 \citep{nextgems2023a,nextgems2023b}. We could
not use Cycle 4 due to unavailability of the necessary variables. The
horizontal resolution of ngc3028 is about 5 km.  Unlike current general
circulation models (GCMs), the strom-resolving version of ICON does not use
convective and cloud parameterisation but relies on explicit simulation of
convection and clouds on the model grid. While this makes the code development
simpler without having to rely on uncertain parameterisations, it can miss
smaller-scale clouds below the grid resolution.  Turbulence and cloud
microphysics are still parameterised in this model. In our analysis, we used
four years of coupled simulation output between years 2021 and 2024
(inclusive). Because the model is free-running, weather and climate
oscillations (such as the El Niño--Southern Oscillation) are not expected to be
equivalent to reality at the same time and place. To compare with observations
taken during different years (2010--2021), we compared the model output with
observations at the same time of year and geographical location.

ICON is a very versatile model, allowing for simulations ranging from
coarse-resolution ESM simulations, GSRM simulations, limited area model (LAM)
simulations, to large eddy simulations (LES), for both weather prediction and
climate projections.

ICON uses the atmospheric component ICON-A \citep{giorgetta2018}, whose
physics is derived from ECHAM6 \citep{stevens2013}, and the ocean component
ICON-O \citep{korn2022}.

Earlier runs of the GSRM ICON from DYAMOND were evaluated by
\cite{mauritsen2022}.

@@

Similar to our results \cite{cesana2022} showed that CMIP6 also tend to
underestimate cloud occurrence above 2 km over the SO, although their analysis
in this case was limited to liquid clouds.
